---
title: "Data_exploration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data set


```{r}
library("tidyverse")
library("glmnet")
```

```{r pressure, echo=FALSE}

mextract = function(model, screen = FALSE, redundant = TRUE){
  # what's the name of the dependent variable?
  yname = deparse(stats::formula(model)[[2]])
  # Set up the data frames for use
  data = stats::model.frame(model)
  X = stats::model.matrix(model)
  n = nrow(X)
  # full model plus redundant variable
  exp.vars = names(model$coefficients)[names(model$coefficients) != "(Intercept)"]

  if (redundant) {
    REDUNDANT.VARIABLE = stats::runif(n, min = 0, max = 1)
    X = cbind(X,REDUNDANT.VARIABLE)
    data = cbind(data,REDUNDANT.VARIABLE)
    exp.vars = c(exp.vars,"REDUNDANT.VARIABLE")
  }
  if (colnames(X)[1] == "(Intercept)") {
    # overwrite intercept with y-variable
    X[,1] = stats::model.frame(model)[,yname]
  } else {
    X = cbind(stats::model.frame(model)[,yname],X)
  }
  colnames(X)[1] = yname
  X = data.frame(X)
  fixed = stats::as.formula(c(paste(yname, "~"),
                       paste(colnames(X)[-1], collapse = "+")))
  Xy = X[c(2:ncol(X),1)]

  k = length(exp.vars) + 1 # +1 for intercept
  if (screen) {
    if (!requireNamespace("mvoutlier", quietly = TRUE)) {
      stop("mvoutlier package needed when screen=TRUE. Please install it.",
           call. = FALSE)
    }
    x.mad = apply(Xy, 2, stats::mad)
    Xy.sub = Xy[,which(x.mad != 0)]
    Xy = Xy[mvoutlier::pcout(Xy.sub)$wfinal01 == 1,]
    n = dim(Xy)[1]
    if (k >= n) {
      warning("Screening deleted too many observations.")
      return()
    }
  }
  wts = model$weights
  if (is.element("glm",class(model))) {
    wts = model$prior.weights
    Xy[,yname] = model$y
  }
  if (is.null(wts)) {
    wts = rep(1,n)
  }

  return(list(yname = yname, fixed = fixed,
              wts = wts, X = Xy, k = k,
              n = n, exp.vars = exp.vars,
              data = data, family = stats::family(model)))
}
```


```{r}
bglmnet = function(mf, nlambda = 100, lambda = NULL, B = 100,
                   penalty.factor, screen = FALSE,
                   redundant = TRUE,
                   cores = NULL,
                   force.in = NULL,
                   seed = NULL) {
   if(!is.null(seed)) {
    set.seed(seed)
  }
  
  m = mextract(mf, screen = screen,
               redundant = redundant)
  
  fixed = m$fixed
  yname = m$yname
  X = m$X
  kf = m$k
  n = m$n
  family = m$family
  fam = family$family
  if (!is.element(fam,c("gaussian", "binomial", "poisson","multinomial", "cox", "mgaussian"))) {
    stop(paste("family is",fam,
               "but it needs to be one of gaussian, binomial, poisson, multinomial, cox, mgaussian"),
         call. = FALSE)
  }
  Xy = m$X
  kf = m$k
  X = Xy[,1:(kf - 1)]
  Y = Xy[,kf]
  n = m$n
  n.obs = n
  X = scale(X) * sqrt(n)/sqrt(n - 1)
  #X[which(is.na(X))] = 0
  X = cbind(1, X)
  colnames(X) = c("(Intercept)",colnames(X)[-1])
  if (missing(penalty.factor)) {
    # link this with force.in
    penalty.factor = c(0, rep(1, kf))
  }
  if (!is.null(lambda)) {
    nlambda = length(lambda)
  }
  temp = glmnet::glmnet(X, Y, alpha = 1, nlambda = nlambda,
                        lambda = lambda,
                        penalty.factor = penalty.factor,
                        weights = m$wts)
  mat = NULL
  # redefine lambda explicitly
  lambda = temp$lambda
  nlambda = length(lambda)
  compteur = matrix(0, kf, nlambda)
  mfstar = do.call("glm",list(fixed, data = Xy, family = family, weights = m$wts))
  #ystar = stats::simulate(object = mfstar, nsim = B)
  #ystar[is.na(ystar)] = Xy[is.na(ystar),yname]
  
  betaboot = array(0,dim = c(kf,nlambda,B))
  rownames(betaboot) = names(mfstar$coef)
  for (j in 1:B) {
    wts = stats::rexp(n = n.obs, rate = 1) * m$wts
    for (i in 1:nlambda) {
      temp = glmnet::glmnet(X, Y, #ystar[,j], 
                            alpha = 1,
                            lambda = lambda[i],
                            intercept = TRUE,
                            #penalty.factor = penalty.factor,
                            family = fam,
                            weights = wts)
      betaboot[,i,j] = (temp$beta[, 1] != 0)
    }
  }
  # looking at model selection across bootstrap replications
  get_unique_mods = function(x) unique(t((x)))
  get.names = function(x) paste(names(x)[x == 1],collapse = "+")
  
  
  prob = NULL # hack for no visible binding for global variable 'prob'
  mod.sum = betaboot %>% 
    apply(3, get_unique_mods) %>% 
    do.call(rbind, .) %>% 
    data.frame() %>% 
    dplyr::mutate(k = rowSums(.)) %>% 
    dplyr::group_by_all() %>% 
    dplyr::summarise(n = n()) %>% 
    dplyr::arrange(k,n) %>% 
    dplyr::mutate(
      k = k + 1
    ) %>% 
    dplyr::group_by(k) %>% 
    dplyr::mutate(
      prob = n/sum(n)
    ) %>% dplyr::ungroup()
  mod.sum$mod.names = mod.sum %>% dplyr::select(-k,-n,-prob) %>% apply(.,1,get.names)
  mod.sum$mod.names[mod.sum$mod.names == ""] = "1"
  mod.sum$mod.names = gsub(pattern = "X.Intercept.",
                           replacement = "1", x = mod.sum$mod.names)
  mod.sum$logLikelihood = NA
  
  for (i in 1:nrow(mod.sum)) {
    # don't need to do this for models that include REDUNDANT.VARIABLE
    mod.sum$logLikelihood[i] = stats::logLik(stats::glm(stats::as.formula(paste(yname,"~",mod.sum$mod.names[i])),
                                                        data = Xy,
                                                        family = family,
                                                        weights = m$wts))
  }
  
  # looking at variable inclusion across lambda values
  compteur2 = apply(betaboot, c(1,2), sum)
  probavariable = compteur2/B
  colnames(probavariable) = round(lambda,3)
  mods = list()
  for (k in 1:length(lambda)) {
    mods[[k]] = table(apply(betaboot[,k,],2,get.names))
  }

  blarout = list(frequency = probavariable,
                 lambda = lambda,
                 mods = mods,
                 mod.sum = mod.sum,
                 screen = screen,
                 vars = names(mfstar$coef),
                 call = match.call())
  class(blarout) = "bglmnet"
  return(blarout)
}

plot.bglmnet = function(x, highlight, interactive = FALSE, 
                        classic = NULL, 
                        tag = NULL, shiny = FALSE,
                        which=c("vip","boot","boot_size"),
                        width=800, height=400, fontSize=12,
                        left=50, top=30,
                        chartWidth="60%",
                        chartHeight="80%",
                        axisTitlesPosition="out",
                        dataOpacity=0.5,
                        options=NULL,
                        hAxis.logScale = TRUE,
                        ylim, text = FALSE,
                        backgroundColor = 'transparent',
                        legend.position = "right",
                        jitterk = 0.1,
                        srt = 45,
                        max.circle = 15,
                        min.prob = 0.1, ...) {
  if (!is.null(classic)) interactive = !classic
  if (backgroundColor == "transparent") {
    backgroundColor = "{stroke:null, fill:'null', strokeSize: 0}"
  } else {
    backgroundColor = paste("{stroke:null, fill:'",backgroundColor,
                            "', strokeSize: 0}", sep = "")
  }
  B = sum(x$mods[[1]])
  gvis.hAxis = paste("{title:'Penalty parameter',
                    logScale:'",hAxis.logScale,"' ,
                    baseline:",0," ,
                     maxValue:",max(x$lambda)*1.1," ,
                     minValue:",min(x$lambda),"}",sep="")
  
  
  if (base::missing(highlight)) {
    no.highlight = TRUE
    highlight =  x$vars[2]
  }
  
  reverselog_trans <- function(base = 10) {
    trans <- function(x) -base::log(x, base)
    inv <- function(x) base^(-x)
    scales::trans_new(base::paste0("reverselog-", base::format(base)), 
                      trans, inv, 
                      scales::log_breaks(base = base), 
                      domain = c(1e-100, Inf))
  }
  
  
  if ("vip" %in% which) {
    
    var.names = x$vars[x$vars != "(Intercept)"]
    p.var = t(x$freq)
    p.var = p.var[,colnames(p.var) %in% var.names]
    sortnames = names(sort(apply(p.var, 2, mean), decreasing = TRUE))
    vip.df = p.var[,sortnames]
    rownames(vip.df) = NULL
    vip.df = data.frame(lambda = x$lambda, vip.df)

    colnames(vip.df) = gsub("REDUNDANT.VARIABLE", "RV", colnames(vip.df))
    sortnames = gsub("REDUNDANT.VARIABLE", "RV", sortnames)
    if(!interactive) { 
      lambda = NULL # hack for no visible binding for global variable 'lambda'
      ggdat = vip.df %>% 
        tidyr::gather(-lambda, key = "variable", value = "prob")
      p = ggplot2::ggplot(
        data = ggdat,
        ggplot2::aes_string(
          x = "lambda",
          y = "prob",
          colour = "variable"
        )
      ) +
        ggplot2::geom_line() +
        ggplot2::theme_bw(base_size = 14) +
        ggplot2::labs(y = "Bootstrap inclusion probability",
                      x = "Penalty") +
        ggplot2::scale_x_log10() +
        ggplot2::theme(
          legend.title = ggplot2::element_blank(),
          legend.key = ggplot2::element_blank(),
          legend.position = legend.position
        )
      
      return(p)
      
    } else { # interactive = TRUE
      
      gvis.title = "Variable inclusion plot (lasso)"
      chartArea = paste("{left:",left,
                        ",top:",top,
                        ",width:'",chartWidth,
                        "',height:'",chartHeight,"'}", sep = "")
      if (is.null(options)) {
        use.options = list(title = gvis.title,
                           fontSize = fontSize,
                           vAxis = "{title:'Bootstrapped probability'}",
                           hAxis = gvis.hAxis,
                           sizeAxis = "{minValue: 0, minSize: 1,
                           maxSize: 20, maxValue:1}",
                           axisTitlesPosition = axisTitlesPosition,
                           chartArea = chartArea,
                           width = width, height = height,
                           backgroundColor = backgroundColor,
                           annotations = "{style:'line'}")
    } else {use.options = options}
      fplot = googleVis::gvisLineChart(data = vip.df,
                                       xvar = "lambda",
                                       yvar = sortnames,
                                       options = use.options)
      if(shiny){
        return(fplot)
      } else {
        return(graphics::plot(fplot, tag = tag))
      }
    }
  }
  
  if("boot" %in% which){
    
    l.vec = rep(x$lambda, times = lapply(x$mods,length))
    mod.vec = unlist(x$mods)
    mod.names = names(mod.vec)
    mod.names[mod.names==""] = "1"
    mod.vec.counts = as.numeric(mod.vec)
    mod.vec.prob = mod.vec.counts/B
    df.temp = data.frame(l.vec,mod.vec.counts,mod.vec.prob,mod.names)
    # remove redundant variables
    df = df.temp[-grep("REDUNDANT.VARIABLE",df.temp$mod.names),]
    df.full = merge(df,x$mod.sum,all.x = TRUE)
    if(all(df.full$mod.vec.prob<min.prob))
      min.prob = stats::quantile(df.full$mod.vec.prob,0.75)
    df.sub = subset(df.full, df.full$mod.vec.prob > min.prob)
    df.sub$mod.names = as.character(df.sub$mod.names)
    
    
    mod.parts = lapply(df.sub$mod.names,FUN = strsplit,"+",fixed=TRUE)
    find.var = function(x,highlight){
      is.element(highlight,unlist(x))
    }
    var.ident = unlist(lapply(mod.parts, find.var,highlight=highlight))
    var.ident[var.ident==TRUE] =  paste("With",highlight)
    var.ident[var.ident==FALSE] =  paste("Without",highlight)
    df.sub$var.ident = var.ident
    df.sub$m2ll = -2*df.sub$logLikelihood
    
    if (! interactive) {
      
      p = ggplot2::ggplot(
        data = df.sub,
        ggplot2::aes_string(x = "l.vec", 
                            y = "m2ll",
                            label = "mod.names")) + 
        ggplot2::geom_jitter(
          ggplot2::aes_string(size = "mod.vec.prob",
                              fill = "var.ident"),
          shape = 21,
          width = 0.0,
          alpha = 0.4) + 
        ggplot2::scale_x_continuous(trans = reverselog_trans()) +
        ggplot2::theme_bw(base_size = 14) + 
        ggplot2::labs(y = "-2*Log-likelihood",
                      x = "Penalty parameter") + 
        ggplot2::theme(legend.title = ggplot2::element_blank(),
                       legend.key = ggplot2::element_blank(),
                       legend.position = "right") +
        ggplot2::scale_fill_manual(values = ggplot2::alpha(c("red","blue"), .4)) +
        ggplot2::guides(
          fill = ggplot2::guide_legend(
            override.aes = list(
              shape = 22,
              size = 5,
              fill = ggplot2::alpha(c("red","blue"), .4)
            )
          )
        )
      
      # if (!missing(ylim)) 
      #   p = p + ggplot2::ylim(ylim[1],ylim[2])
      
      return(p)
      
    } else { # interactive = TRUE
      
      gvis.title = paste("Model stability plot for glmnet",sep="")
      #x.ticks=paste(1:max(x$lk$k),collapse=",")
      chartArea = paste("{left:",left,
                        ",top:",top,
                        ",width:'",chartWidth,
                        "',height:'",chartHeight,"'}",sep="")
      bubble = paste("{opacity:",dataOpacity,
                     ", textStyle: {color: 'none'}}",sep="")
      
      if(is.null(options)){
        use.options=list(title=gvis.title,
                         fontSize = fontSize,
                         vAxis="{title:'-2*Log-likelihood'}",
                         hAxis=gvis.hAxis,
                         sizeAxis = "{minValue: 0, minSize: 1,
                         maxSize: 20, maxValue:1}",
                         axisTitlesPosition=axisTitlesPosition,
                         bubble = bubble,
                         chartArea=chartArea,
                         width=width, height=height,
                         backgroundColor=backgroundColor,
                         explorer= "{axis: 'vertical',
                         keepInBounds: true,
                         maxZoomOut: 1,
                         maxZoomIn: 0.01,
                         actions: ['dragToZoom',
                         'rightClickToReset']}")
      } else {use.options = options}
      
      fplot = googleVis::gvisBubbleChart(data = df.sub,
                                         idvar = "mod.names",
                                         xvar = "l.vec",
                                         yvar = "m2ll", 
                                         colorvar = "var.ident",
                                         sizevar = "mod.vec.prob",
                                         options = use.options)
      if(shiny){
        return(fplot)
      } else {
        graphics::plot(fplot, tag = tag)
      }
    }
  }
  
  if("boot_size" %in% which) {
    
    
    pd = x$mod.sum
    pd$var.ident.tf = NA
    pd$var.ident.tf = pd[, highlight[1]] == 1
    #vi = var.ident
    pd$var.ident = NA
    pd$var.ident[pd$var.ident.tf == TRUE] = paste("With", highlight[1])
    pd$var.ident[pd$var.ident.tf == FALSE] = paste("Without", highlight[1])
    pd$m2ll = -2*pd$logLikelihood
    
    if(!interactive){
      
      p = ggplot2::ggplot(
        pd,
        ggplot2::aes_string(
          x = "k",
          y = "m2ll",
          group = "var.ident",
          label = "mod.names"
        )) +
        ggplot2::geom_jitter(
          ggplot2::aes_string(size = "prob",
                              fill = "var.ident"),
          shape = 21,
          width = jitterk
        ) +
        ggplot2::scale_size(range = c(0, max.circle)) +
        ggplot2::theme_bw(base_size = 14) +
        ggplot2::ylab("-2*Log-likelihood") +
        ggplot2::xlab("Number of parameters") +
        ggplot2::theme(
          legend.title = ggplot2::element_blank(),
          legend.key = ggplot2::element_blank(),
          legend.position = legend.position
        ) +
        ggplot2::scale_fill_manual(values = ggplot2::alpha(c("red", "blue"), .4)) +
        ggplot2::guides(fill = ggplot2::guide_legend(override.aes = list(
          shape = 22,
          size = 5,
          fill = ggplot2::alpha(c("red", "blue"), .4)
        )))
      if (!base::missing(ylim))
        p = p + ggplot2::ylim(ylim[1], ylim[2])
      if (text) {
        p = p + ggplot2::geom_text(hjust = 0, angle = srt)
      }
      return(p)  
      
    } else { # interactive = TRUE
      gvis.title = paste("Model stability plot (lasso)", sep = "")
      x.ticks = paste(1:max(pd$k), collapse = ",")
      gvis.hAxis = paste(
        "{title:'Number of parameters',
        maxValue:",
        max(pd$k) + 0.5,
        " ,
        minValue:",
        0.5,
        " ,
        ticks: [",
        x.ticks,
        "]}"
      )
      y.min = min(pd$m2ll) - 0.02*(max(pd$m2ll)- min(pd$m2ll))
      gvis.vAxis = paste("{title:'-2*Log-likelihood', minValue:",
                         y.min, "}")
      chartArea = paste(
        "{left:", left, ",top:", top,
        ",width:'", chartWidth, "',height:'", chartHeight,
        "'}", sep = ""
      )
      bubble = paste("{opacity:",
                     dataOpacity,
                     ", textStyle: {color: 'none'}}",
                     sep = "")
      if (is.null(options)) {
        use.options = list(
          title = gvis.title,
          fontSize = fontSize,
          vAxis = gvis.vAxis,
          hAxis = gvis.hAxis,
          sizeAxis = "{minValue: 0, minSize: 1,
          maxSize: 20, maxValue:1}",
          axisTitlesPosition = axisTitlesPosition,
          bubble = bubble,
          chartArea = chartArea,
          width = width,
          height = height,
          backgroundColor = backgroundColor,
          explorer = "{axis: 'vertical',
          keepInBounds: true,
          maxZoomOut: 1,
          maxZoomIn: 0.01,
          actions: ['dragToZoom',
          'rightClickToReset']}"
        )
      } else {
        use.options = options
      }
      fplot = googleVis::gvisBubbleChart(
        data = pd,
        idvar = "mod.names",
        xvar = "k",
        yvar = "m2ll",
        colorvar = "var.ident",
        sizevar = "prob",
        options = use.options
      )
      if (shiny) {
        return(fplot)
      } else {
        graphics::plot(fplot, tag = tag)
      }
    }
  } else return(invisible())
  
}
```

```{r}
load("abaloneClean.rda")
```

```{r}
predictors = colnames(abalone %>% select(-Sex, -Age))
weights = c("WholeWeight", "ShellWeight", "ShuckedWeight", "VisceraWeight")
abalone = abalone %>% mutate(Sex = factor(Sex))
```

```{r}
# skimr::skim(abalone)
```

# Assess normality of plots
```{r}
for(c in predictors)
{
  a = ggplot(abalone, aes_string(x=c)) + geom_histogram(bins=30)
  b = ggplot(abalone, aes_string(y=c)) + geom_boxplot()
  p = ggplot(abalone, aes_string(sample=c)) + stat_qq() + stat_qq_line()
  gridExtra::grid.arrange(a, b, p, nrow=1)
}
print(ggplot(abalone, aes(x=Sex)) + geom_bar())
```

Most distributions are skewed, some outliers on qqplots of height.

# Assess whether male/female/infants have similar distributions of predictors/response
```{r}
for(c in predictors)
{
  print(c)
  eq = paste(c, "~Sex", sep="")
  av = aov(as.formula(eq), abalone)
  print(summary(av))
  
  print(ggplot(abalone, aes_string(y=c, x="Sex")) + geom_boxplot())
}
```

```{r}
abalone_adult = abalone %>% filter(Sex != "Infant")
```

```{r}
for(c in predictors)
{
  eq = paste(c, "~Sex", sep="")
  print(t.test(as.formula(eq), abalone_adult))
  print(ggplot(abalone_adult, aes_string(y=c, x="Sex")) + geom_boxplot())
}
```

The different populations have very different predictor/response distributions.

# Linear models
```{r}
abalone.male = abalone %>% 
  filter(Sex == "Male") %>%
  select(-Sex)
abalone.female = abalone %>% 
  filter(Sex == "Female") %>%
  select(-Sex)
abalone.infant = abalone %>% 
  filter(Sex == "Infant") %>%
  select(-Sex)
```

Proposed models:
```{r}
models = c(
  "Rings~Length+Diameter+Height",
  "Rings~WholeWeight",
  "Rings~Length+Diameter+Height+WholeWeight",
  "Rings~Length+Diameter+Height+ShellWeight+ShuckedWeight",
  "Rings~ShellWeight+ShuckedWeight",
  "Rings~Length+Diameter+Height+ShellWeight+ShuckedWeight+VisceraWeight",
  "Rings~ShellWeight+ShuckedWeight+VisceraWeight",
  "Rings~Length+Diameter+Height+ShuckedWeight+VisceraWeight",
  "Rings~ShuckedWeight+VisceraWeight",
  "Rings~Length+Diameter+Height+ShellWeight+VisceraWeight",
  "Rings~ShellWeight+VisceraWeight"
)
```

Pairwise plots:
```{r}
pairs = matrix(c(
  "Length", "Diameter",
  "Height", "Length",
  "Diameter", "Height",
  "WholeWeight", "Length",
  "WholeWeight", "Height",
  "WholeWeight", "Diameter",
  "ShuckedWeight", "VisceraWeight",
  "ShuckedWeight", "Length",
  "ShuckedWeight", "Height",
  "ShuckedWeight", "Diameter",
  "VisceraWeight", "Length",
  "VisceraWeight", "Height",
  "VisceraWeight", "Diameter"
  ), ncol=2, byrow=T
)
```

## Male
```{r}
lm.male = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.male)
  print(summary(lm_result))
  plot(lm_result)
  lm.male[[i]] = lm_result
}
```

Diagnostic plots suggest 531 is an outlier. 

```{r}
# Removing the outlier point:
abalone.male = abalone.male[-531, ]
```
```{r}
lm.male = NULL
for(i in 1:length(models))
{
  lm_result = lm(models[i], abalone.male)
  print(summary(lm_result))
  plot(lm_result)
  lm.male[[i]] = lm_result
}
```

May be a logarithmic relationship between rings and weight.

```{r}
for(i in 1:nrow(pairs))
{
  x = pairs[i, 1]
  z = pairs[i, 2]
  p = ggplot(abalone.male, aes_string(x=x, y="Rings", colour=z)) +
      geom_point() +
      geom_smooth(method="lm") + 
      labs(title=paste("No. rings versus.", x, "and", z, "(colour)"))
  print(p)
}
```

```{r}
for(w in weights)
{
  abalone.male[[paste(w, "_exp", sep="")]] = exp(abalone.male[[w]])
}
```

New proposed models:
```{r}
models_exp = c(
  "Rings~Length+Diameter+Height",
  "Rings~WholeWeight_exp",
  "Rings~Length+Diameter+Height+WholeWeight_exp",
  "Rings~Length+Diameter+Height+ShellWeight_exp+ShuckedWeight_exp",
  "Rings~ShellWeight_exp+ShuckedWeight_exp",
  "Rings~Length+Diameter+Height+ShellWeight_exp+ShuckedWeight_exp+VisceraWeight_exp",
  "Rings~ShellWeight_exp+ShuckedWeight_exp+VisceraWeight_exp",
  "Rings~Length+Diameter+Height+ShuckedWeight_exp+VisceraWeight_exp",
  "Rings~ShuckedWeight_exp+VisceraWeight_exp",
  "Rings~Length+Diameter+Height+ShellWeight_exp+VisceraWeight_exp",
  "Rings~ShellWeight_exp+VisceraWeight_exp"
)
```

New pairwise plots:
```{r}
pairs_exp = matrix(c(
  "Length", "Diameter",
  "Height", "Length",
  "Diameter", "Height",
  "WholeWeight_exp", "Length",
  "WholeWeight_exp", "Height",
  "WholeWeight_exp", "Diameter",
  "ShuckedWeight_exp", "VisceraWeight_exp",
  "ShuckedWeight_exp", "Length",
  "ShuckedWeight_exp", "Height",
  "ShuckedWeight_exp", "Diameter",
  "VisceraWeight_exp", "Length",
  "VisceraWeight_exp", "Height",
  "VisceraWeight_exp", "Diameter"
  ), ncol=2, byrow=T
)
```

```{r}
lm.male_exp = NULL
for(i in 1:length(models_exp))
{
  lm_result = lm(models_exp[i], abalone.male)
  print(summary(lm_result))
  plot(lm_result)
  lm.male_exp[[i]] = lm_result
}
```

```{r}
for(i in 1:nrow(pairs_exp))
{
  x = pairs_exp[i, 1]
  z = pairs_exp[i, 2]
  p = ggplot(abalone.male, aes_string(x=x, y="Rings", colour=z)) +
      geom_point() +
      geom_smooth(method="lm") + 
      labs(title=paste("No. rings versus.", x, "and", z, "(colour)"))
  print(p)
}
```

```{r}
for(i in 1:length(lm.male))
{
  print(i)
  a = summary(lm.male[[i]])
  b = summary(lm.male_exp[[i]])
  print(paste("MODEL A:   ", a$sigma, a$r.squared, a$adj.r.squared, sep=" "))
  print(paste("MODEL B:   ", b$sigma, b$r.squared, b$adj.r.squared, sep=" "))
  print("")
}
```

Model 4 is the best model but we can drop some variables. First drop length:
```{r}
best.male = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.male)
summary(best.male)
plot(bglmnet(best.male, seed = 1, B=100))
```

This is the final model.

```{r}
abalone.male = abalone.male %>% select(all_of(colnames(abalone.female)))
```

## Female
Assume that there is no log relationship (as we did not observe one in the males). Then we just have to test model assumptions on the proposed list.
```{r}
lm.female = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.female)
  print(summary(lm_result))
  plot(lm_result)
  lm.female[[i]] = lm_result
}
```

Diagnostic plots suggest 621 is an outlier. 

```{r}
# Removing the outlier point:
abalone.female = abalone.female[-621, ]
```

```{r}
lm.female = NULL
for(i in 1:length(models))
{
  lm_result = lm(models[i], abalone.female)
  print(summary(lm_result))
  plot(lm_result)
  lm.female[[i]] = lm_result
}
```

```{r}
for(i in 1:nrow(pairs))
{
  x = pairs[i, 1]
  z = pairs[i, 2]
  p = ggplot(abalone.female, aes_string(x=x, y="Rings", colour=z)) +
      geom_point() +
      geom_smooth(method="lm") + 
      labs(title=paste("No. rings versus.", x, "and", z, "(colour)"))
  print(p)
}
```

```{r}
for(i in 1:length(lm.female))
{
  print(i)
  a = summary(lm.female[[i]])
  print(paste("MODEL:   ", a$sigma, a$r.squared, a$adj.r.squared, sep=" "))
  print("")
}
```

Model 4/6 are the best models.

Model 4 with length dropped:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.female)
summary(res)
plot(bglmnet(res, seed = 1, B=100))
```
Drop diameter too:
```{r}
res = lm(Rings~Height+ShellWeight+ShuckedWeight, abalone.female)
summary(res)
plot(bglmnet(res, seed = 1, B=100))
```

Model 6 with VisceraWeight dropped:
```{r}
best.female = lm(Rings~Length+Diameter+Height+ShellWeight+ShuckedWeight, abalone.female)
summary(best.female)
plot(bglmnet(best.female, seed = 1, B=100))
```

This is the best model for the female population.

## Infant
Assume that there is no log relationship (as we did not observe one in the males). Then we just have to test model assumptions on the proposed list.
```{r}
lm.infant = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.infant)
  print(summary(lm_result))
  plot(lm_result)
  lm.infant[[i]] = lm_result
}
```

Diagnostic plots suggest 312 is an outlier. 

```{r}
# Removing the outlier point:
abalone.infant = abalone.infant[-312, ]
```

```{r}
lm.infant = NULL
for(i in 1:length(models))
{
  lm_result = lm(models[i], abalone.infant)
  print(summary(lm_result))
  plot(lm_result)
  lm.infant[[i]] = lm_result
}
```

1246 may also be an outlier
```{r}
# Removing the outlier point:
abalone.infant = abalone.infant[-1246, ]
```

```{r}
lm.infant = NULL
for(i in 1:length(models))
{
  lm_result = lm(models[i], abalone.infant)
  print(summary(lm_result))
  plot(lm_result)
  lm.infant[[i]] = lm_result
}
```

```{r}
for(i in 1:nrow(pairs))
{
  x = pairs[i, 1]
  z = pairs[i, 2]
  p = ggplot(abalone.infant, aes_string(x=x, y="Rings", colour=z)) +
      geom_point() +
      geom_smooth(method="lm") + 
      labs(title=paste("No. rings versus.", x, "and", z, "(colour)"))
  print(p)
}
```

```{r}
for(i in 1:length(lm.infant))
{
  print(i)
  a = summary(lm.infant[[i]])
  print(paste("MODEL:   ", a$sigma, a$r.squared, a$adj.r.squared, sep=" "))
}
```

Model 4/6 are the best models.

Model 4 with length dropped:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.infant)
summary(res)
plot(bglmnet(res, seed = 1, B=100))
```

Drop diameter too:
```{r}
res = lm(Rings~Height+ShellWeight+ShuckedWeight, abalone.infant)
summary(res)
plot(bglmnet(res, seed = 1, B=100))
```

Model 6 with VisceraWeight dropped:
```{r}
res = lm(Rings~Length+Diameter+Height+ShellWeight+ShuckedWeight, abalone.infant)
summary(res)
plot(bglmnet(res, seed = 1, B=100))
```

Model 6 with Length dropped:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.infant)
summary(res)
plot(bglmnet(res, seed = 1, B=100))
```

Model 6 with Diameter dropped:
```{r}
best.infant = lm(Rings~Height+ShellWeight+ShuckedWeight, abalone.infant)
summary(best.infant)
plot(bglmnet(res, seed = 1, B=100))
```

This is the same as Model 4 following variable selection, and is the best model for infants.


## Adults
```{r}
abalone.adults = rbind(abalone.male, abalone.female)
```

```{r}
lm.adults = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.adults)
  print(summary(lm_result))
  plot(lm_result)
  lm.adults[[i]] = lm_result
}
```

```{r}
abalone.adults = abalone.adults[-2148,]
```

```{r}
lm.adults = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.adults)
  print(summary(lm_result))
  plot(lm_result)
  lm.adults[[i]] = lm_result
}
```

```{r}
abalone.adults = abalone.adults[-531,]
```

```{r}
lm.adults = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.adults)
  print(summary(lm_result))
  plot(lm_result)
  lm.adults[[i]] = lm_result
}
```

```{r}
for(i in 1:length(lm.adults))
{
  print(i)
  a = summary(lm.adults[[i]])
  print(paste("MODEL:   ", a$sigma, a$r.squared, a$adj.r.squared, sep=" "))
}
```

Model 4 without Length:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.adults)
summary(res)
plot(bglmnet(res, seed = 1, B=100))
```

Model 6 without Length:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight+VisceraWeight, abalone.adults)
summary(res)
plot(bglmnet(res, seed = 1, B=100))
```

Model 6 without VisceraWeight:
```{r}
best.adults = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.adults)
summary(best.adults)
plot(bglmnet(best.adults, seed = 1, B=10))


```

Best model for adults.

## Entire population
```{r}
abalone.full = rbind(abalone.male, abalone.female, abalone.infant)
```

```{r}
lm.full = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.full)
  print(summary(lm_result))
  plot(lm_result)
  lm.full[[i]] = lm_result
}
```

```{r}
abalone.full = abalone.full[-2148,]
```

```{r}
lm.full = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.full)
  print(summary(lm_result))
  plot(lm_result)
  lm.full[[i]] = lm_result
}
```

```{r}
abalone.full = abalone.full[-531,]
```

```{r}
lm.full = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.full)
  print(summary(lm_result))
  plot(lm_result)
  lm.full[[i]] = lm_result
}
```

```{r}
for(i in 1:length(lm.full))
{
  print(i)
  a = summary(lm.full[[i]])
  print(paste("MODEL:   ", a$sigma, a$r.squared, a$adj.r.squared, sep=" "))
}
```

Model 4 without Length:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.full)
summary(res)
plot(bglmnet(res, seed = 1, B=10))
```

Model 6 without VisceraWeight:
```{r}
res = lm(Rings~Length+Diameter+Height+ShellWeight+ShuckedWeight, abalone.full)
summary(res)
plot(bglmnet(res, seed = 1, B=10))
```

Model 6 without Length:
```{r}
best.full = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.full)
summary(best.full)
plot(bglmnet(best.full, seed = 1, B=10))
```

```{r}
summary(lm(Rings~Diameter+Height+ShuckedWeight, abalone.full))
```