---
title: "Data_exploration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data set

### Inspection

```{r}
library("tidyverse")

abalone = read.csv("abalone.data",
                   col.names = c("Sex", "Length", "Diameter", "Height", "WholeWeight", "ShuckedWeight", "VisceraWeight", "ShellWeight", "Rings"))

abalone = abalone %>% mutate(
  Sex = case_when(
    Sex == "M" ~ "Male",
    Sex == "F"~ "Female",
    Sex == "I" ~ "Infant"
  ),
  Age = Rings + 1.5
)

head(abalone)
```

```{r}
abalone %>% nrow()
```

There is 4176 rows in total.

### Data cleansing

After comparing these `*Weight` columns for a while, I found a few anomaly data:

```{r}
abalone[(abalone$WholeWeight < abalone$ShuckedWeight), ]
```

These four rows has the weight of the whole abalone less than the weight of their meat that is likely to be a typo in the data.

I also found a similar anomaly:

```{r}
abalone[(abalone$WholeWeight < abalone$ShellWeight), ]
```

The whole weight of this row is less than its weight after it's being dried, which is extremely unlikely to happen, if not impossible.

I also found two rows that has their `Height` equals to 0:

```{r}
abalone[abalone$Height <= 0, ]
```

The second row '3996' is also the one that has its whole weight less than its shell weight.

To obtain a useful database, these anomalies are removed from the data frame:

```{r}
abalone = abalone %>% filter(WholeWeight > ShuckedWeight, WholeWeight > ShellWeight, Height > 0)
save(abalone, file="abaloneClean.rda")
nrow(abalone)
```

After the data cleansing, 6 rows are removed in total.

### Explore correlation matrix

```{r}
library("qtlcharts")

qtlcharts::iplotCorr(abalone[, c("Length", "Diameter", "Height", "WholeWeight", "ShuckedWeight", "VisceraWeight", "ShellWeight", "Age")])
```

After exploring the correlation matrix above, it seems that no variable has a linear relationship with `Age`.

Thus it is necessary to try `log(Age)`:

```{r}
abalone$lAge = log(abalone$Age)
head(abalone)
```

Now try the correlation matrix again:

```{r}
qtlcharts::iplotCorr(abalone[, c("Length", "Diameter", "Height", "WholeWeight", "ShuckedWeight", "VisceraWeight", "ShellWeight", "lAge")])
```

However, it is a possible that for each `Sex`, there is a relationship between `Age` or `log(Age)` to other variable, so lets try again:

For male:

```{r}
qtlcharts::iplotCorr(abalone[abalone$Sex == "Male",
                             c("Length", "Diameter", "Height", "WholeWeight", "ShuckedWeight", "VisceraWeight", "ShellWeight", "Age", "lAge")])
```

For female:

```{r}
qtlcharts::iplotCorr(abalone[abalone$Sex == "Female",
                             c("Length", "Diameter", "Height", "WholeWeight", "ShuckedWeight", "VisceraWeight", "ShellWeight", "Age", "lAge")])
```

For infant:

```{r}
qtlcharts::iplotCorr(abalone[abalone$Sex == "Infant",
                             c("Length", "Diameter", "Height", "WholeWeight", "ShuckedWeight", "VisceraWeight", "ShellWeight", "Age", "lAge")])
```

Unfortunately, in none of the `Sex` in `abalone` does `Age` or `lAge` have any linear relationship with any other variable.

Thus, we have to explore other relationship.M

