---
title: "Data_exploration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data set


```{r}
library("tidyverse")
```

```{r}
load("abaloneClean.rda")
```

```{r}
predictors = colnames(abalone %>% select(-Sex, -Age))
weights = c("WholeWeight", "ShellWeight", "ShuckedWeight", "VisceraWeight")
abalone = abalone %>% mutate(Sex = factor(Sex))
```

```{r}
# skimr::skim(abalone)
```

# Assess normality of plots
```{r}
for(c in predictors)
{
  a = ggplot(abalone, aes_string(x=c)) + geom_histogram(bins=30)
  b = ggplot(abalone, aes_string(y=c)) + geom_boxplot()
  p = ggplot(abalone, aes_string(sample=c)) + stat_qq() + stat_qq_line()
  gridExtra::grid.arrange(a, b, p, nrow=1)
}

print(ggplot(abalone, aes(x=Sex)) + geom_bar())
```

Most distributions are skewed, some outliers on qqplots of height.

# Assess whether male/female/infants have similar distributions of predictors/response
```{r}
for(c in predictors)
{
  print(c)
  eq = paste(c, "~Sex", sep="")
  av = aov(as.formula(eq), abalone)
  print(summary(av))
  
  print(ggplot(abalone, aes_string(y=c, x="Sex")) + geom_boxplot())
}
```

```{r}
abalone_adult = abalone %>% filter(Sex != "Infant")
```

```{r}
for(c in predictors)
{
  eq = paste(c, "~Sex", sep="")
  print(t.test(as.formula(eq), abalone_adult))

  print(ggplot(abalone_adult, aes_string(y=c, x="Sex")) + geom_boxplot())
}
```

The different populations have very different predictor/response distributions.

# Linear models
```{r}
abalone.male = abalone %>% 
  filter(Sex == "Male") %>%
  select(-Sex)
abalone.female = abalone %>% 
  filter(Sex == "Female") %>%
  select(-Sex)
abalone.infant = abalone %>% 
  filter(Sex == "Infant") %>%
  select(-Sex)
```

Proposed models:
```{r}
models = c(
  "Rings~Length+Diameter+Height",
  "Rings~WholeWeight",
  "Rings~Length+Diameter+Height+WholeWeight",
  "Rings~Length+Diameter+Height+ShellWeight+ShuckedWeight",
  "Rings~ShellWeight+ShuckedWeight",
  "Rings~Length+Diameter+Height+ShellWeight+ShuckedWeight+VisceraWeight",
  "Rings~ShellWeight+ShuckedWeight+VisceraWeight",
  "Rings~Length+Diameter+Height+ShuckedWeight+VisceraWeight",
  "Rings~ShuckedWeight+VisceraWeight",
  "Rings~Length+Diameter+Height+ShellWeight+VisceraWeight",
  "Rings~ShellWeight+VisceraWeight"
)
```

Pairwise plots:
```{r}
pairs = matrix(c(
  "Length", "Diameter",
  "Height", "Length",
  "Diameter", "Height",
  "WholeWeight", "Length",
  "WholeWeight", "Height",
  "WholeWeight", "Diameter",
  "ShuckedWeight", "VisceraWeight",
  "ShuckedWeight", "Length",
  "ShuckedWeight", "Height",
  "ShuckedWeight", "Diameter",
  "VisceraWeight", "Length",
  "VisceraWeight", "Height",
  "VisceraWeight", "Diameter"
  ), ncol=2, byrow=T
)
```

## Male
```{r}
lm.male = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.male)
  print(summary(lm_result))
  plot(lm_result)
  lm.male[[i]] = lm_result
}
```

Diagnostic plots suggest 531 is an outlier. 

```{r}
# Removing the outlier point:
abalone.male = abalone.male[-531, ]
```
```{r}
lm.male = NULL
for(i in 1:length(models))
{
  lm_result = lm(models[i], abalone.male)
  print(summary(lm_result))
  plot(lm_result)
  lm.male[[i]] = lm_result
}
```

May be a logarithmic relationship between rings and weight.

```{r}
for(i in 1:nrow(pairs))
{
  x = pairs[i, 1]
  z = pairs[i, 2]
  p = ggplot(abalone.male, aes_string(x=x, y="Rings", colour=z)) +
      geom_point() +
      geom_smooth(method="lm") + 
      labs(title=paste("No. rings versus.", x, "and", z, "(colour)"))
  print(p)
}
```

```{r}
for(w in weights)
{
  abalone.male[[paste(w, "_exp", sep="")]] = exp(abalone.male[[w]])
}
```

New proposed models:
```{r}
models_exp = c(
  "Rings~Length+Diameter+Height",
  "Rings~WholeWeight_exp",
  "Rings~Length+Diameter+Height+WholeWeight_exp",
  "Rings~Length+Diameter+Height+ShellWeight_exp+ShuckedWeight_exp",
  "Rings~ShellWeight_exp+ShuckedWeight_exp",
  "Rings~Length+Diameter+Height+ShellWeight_exp+ShuckedWeight_exp+VisceraWeight_exp",
  "Rings~ShellWeight_exp+ShuckedWeight_exp+VisceraWeight_exp",
  "Rings~Length+Diameter+Height+ShuckedWeight_exp+VisceraWeight_exp",
  "Rings~ShuckedWeight_exp+VisceraWeight_exp",
  "Rings~Length+Diameter+Height+ShellWeight_exp+VisceraWeight_exp",
  "Rings~ShellWeight_exp+VisceraWeight_exp"
)
```

New pairwise plots:
```{r}
pairs_exp = matrix(c(
  "Length", "Diameter",
  "Height", "Length",
  "Diameter", "Height",
  "WholeWeight_exp", "Length",
  "WholeWeight_exp", "Height",
  "WholeWeight_exp", "Diameter",
  "ShuckedWeight_exp", "VisceraWeight_exp",
  "ShuckedWeight_exp", "Length",
  "ShuckedWeight_exp", "Height",
  "ShuckedWeight_exp", "Diameter",
  "VisceraWeight_exp", "Length",
  "VisceraWeight_exp", "Height",
  "VisceraWeight_exp", "Diameter"
  ), ncol=2, byrow=T
)
```

```{r}
lm.male_exp = NULL
for(i in 1:length(models_exp))
{
  lm_result = lm(models_exp[i], abalone.male)
  print(summary(lm_result))
  plot(lm_result)
  lm.male_exp[[i]] = lm_result
}
```

```{r}
for(i in 1:nrow(pairs_exp))
{
  x = pairs_exp[i, 1]
  z = pairs_exp[i, 2]
  p = ggplot(abalone.male, aes_string(x=x, y="Rings", colour=z)) +
      geom_point() +
      geom_smooth(method="lm") + 
      labs(title=paste("No. rings versus.", x, "and", z, "(colour)"))
  print(p)
}
```

```{r}
for(i in 1:length(lm.male))
{
  print(i)
  a = summary(lm.male[[i]])
  b = summary(lm.male_exp[[i]])
  print(paste("MODEL A:   ", a$sigma, a$r.squared, a$adj.r.squared, sep=" "))
  print(paste("MODEL B:   ", b$sigma, b$r.squared, b$adj.r.squared, sep=" "))
  print("")
}
```

Model 4 is the best model but we can drop some variables. First drop length:
```{r}
best.male = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.male)
summary(best.male)
```

This is the final model.

```{r}
abalone.male = abalone.male %>% select(all_of(colnames(abalone.female)))
```

## Female
Assume that there is no log relationship (as we did not observe one in the males). Then we just have to test model assumptions on the proposed list.
```{r}
lm.female = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.female)
  print(summary(lm_result))
  plot(lm_result)
  lm.female[[i]] = lm_result
}
```

Diagnostic plots suggest 621 is an outlier. 

```{r}
# Removing the outlier point:
abalone.female = abalone.female[-621, ]
```

```{r}
lm.female = NULL
for(i in 1:length(models))
{
  lm_result = lm(models[i], abalone.female)
  print(summary(lm_result))
  plot(lm_result)
  lm.female[[i]] = lm_result
}
```

```{r}
for(i in 1:nrow(pairs))
{
  x = pairs[i, 1]
  z = pairs[i, 2]
  p = ggplot(abalone.female, aes_string(x=x, y="Rings", colour=z)) +
      geom_point() +
      geom_smooth(method="lm") + 
      labs(title=paste("No. rings versus.", x, "and", z, "(colour)"))
  print(p)
}
```

```{r}
for(i in 1:length(lm.female))
{
  print(i)
  a = summary(lm.female[[i]])
  print(paste("MODEL:   ", a$sigma, a$r.squared, a$adj.r.squared, sep=" "))
  print("")
}
```

Model 4/6 are the best models.

Model 4 with length dropped:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.female)
summary(res)
```
Drop diameter too:
```{r}
res = lm(Rings~Height+ShellWeight+ShuckedWeight, abalone.female)
summary(res)
```

Model 6 with VisceraWeight dropped:
```{r}
best.female = lm(Rings~Length+Diameter+Height+ShellWeight+ShuckedWeight, abalone.female)
summary(best.female)
```

This is the best model for the female population.

## Infant
Assume that there is no log relationship (as we did not observe one in the males). Then we just have to test model assumptions on the proposed list.
```{r}
lm.infant = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.infant)
  print(summary(lm_result))
  plot(lm_result)
  lm.infant[[i]] = lm_result
}
```

Diagnostic plots suggest 312 is an outlier. 

```{r}
# Removing the outlier point:
abalone.infant = abalone.infant[-312, ]
```

```{r}
lm.infant = NULL
for(i in 1:length(models))
{
  lm_result = lm(models[i], abalone.infant)
  print(summary(lm_result))
  plot(lm_result)
  lm.infant[[i]] = lm_result
}
```

1246 may also be an outlier
```{r}
# Removing the outlier point:
abalone.infant = abalone.infant[-1246, ]
```

```{r}
lm.infant = NULL
for(i in 1:length(models))
{
  lm_result = lm(models[i], abalone.infant)
  print(summary(lm_result))
  plot(lm_result)
  lm.infant[[i]] = lm_result
}
```

```{r}
for(i in 1:nrow(pairs))
{
  x = pairs[i, 1]
  z = pairs[i, 2]
  p = ggplot(abalone.infant, aes_string(x=x, y="Rings", colour=z)) +
      geom_point() +
      geom_smooth(method="lm") + 
      labs(title=paste("No. rings versus.", x, "and", z, "(colour)"))
  print(p)
}
```

```{r}
for(i in 1:length(lm.infant))
{
  print(i)
  a = summary(lm.infant[[i]])
  print(paste("MODEL:   ", a$sigma, a$r.squared, a$adj.r.squared, sep=" "))
}
```

Model 4/6 are the best models.

Model 4 with length dropped:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.infant)
summary(res)
```

Drop diameter too:
```{r}
res = lm(Rings~Height+ShellWeight+ShuckedWeight, abalone.infant)
summary(res)
```

Model 6 with VisceraWeight dropped:
```{r}
res = lm(Rings~Length+Diameter+Height+ShellWeight+ShuckedWeight, abalone.infant)
summary(res)
```

Model 6 with Length dropped:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.infant)
summary(res)
```

Model 6 with Diameter dropped:
```{r}
best.infant = lm(Rings~Height+ShellWeight+ShuckedWeight, abalone.infant)
summary(best.infant)
```

This is the same as Model 4 following variable selection, and is the best model for infants.


## Adults
```{r}
abalone.adults = rbind(abalone.male, abalone.female)
```

```{r}
lm.adults = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.adults)
  print(summary(lm_result))
  plot(lm_result)
  lm.adults[[i]] = lm_result
}
```

```{r}
abalone.adults = abalone.adults[-2148,]
```

```{r}
lm.adults = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.adults)
  print(summary(lm_result))
  plot(lm_result)
  lm.adults[[i]] = lm_result
}
```

```{r}
abalone.adults = abalone.adults[-531,]
```

```{r}
lm.adults = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.adults)
  print(summary(lm_result))
  plot(lm_result)
  lm.adults[[i]] = lm_result
}
```

```{r}
for(i in 1:length(lm.adults))
{
  print(i)
  a = summary(lm.adults[[i]])
  print(paste("MODEL:   ", a$sigma, a$r.squared, a$adj.r.squared, sep=" "))
}
```

Model 4 without Length:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.adults)
summary(res)
```

Model 6 without Length:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight+VisceraWeight, abalone.adults)
summary(res)
```

Model 6 without VisceraWeight:
```{r}
best.adults = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.adults)
summary(best.adults)
```

Best model for adults.

## Entire population
```{r}
abalone.full = rbind(abalone.male, abalone.female, abalone.infant)
```

```{r}
lm.full = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.full)
  print(summary(lm_result))
  plot(lm_result)
  lm.full[[i]] = lm_result
}
```

```{r}
abalone.full = abalone.full[-2148,]
```

```{r}
lm.full = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.full)
  print(summary(lm_result))
  plot(lm_result)
  lm.full[[i]] = lm_result
}
```

```{r}
abalone.full = abalone.full[-531,]
```

```{r}
lm.full = NULL
for(i in 1:length(models))
{
  print(i)
  lm_result = lm(models[i], abalone.full)
  print(summary(lm_result))
  plot(lm_result)
  lm.full[[i]] = lm_result
}
```

```{r}
for(i in 1:length(lm.full))
{
  print(i)
  a = summary(lm.full[[i]])
  print(paste("MODEL:   ", a$sigma, a$r.squared, a$adj.r.squared, sep=" "))
}
```

Model 4 without Length:
```{r}
res = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.full)
summary(res)
```

Model 6 without VisceraWeight:
```{r}
res = lm(Rings~Length+Diameter+Height+ShellWeight+ShuckedWeight, abalone.full)
summary(res)
```

Model 6 without Length:
```{r}
best.full = lm(Rings~Diameter+Height+ShellWeight+ShuckedWeight, abalone.full)
summary(best.full)
```

```{r}
summary(lm(Rings~Diameter+Height+ShuckedWeight, abalone.full))
```

